### 2016-03-14 paper/icfp16-submission/introduction.tex
emphasize jump-compression in intro

\begin{figure}[t]
\centering
\begin{minipage}{.49\linewidth}
\centering
\begin{code}
  let rec fac n =
    if n <= 0 then
      true
    else
      n * fac (n-1)
\end{code}
\end{minipage}
% \hspace{0.3in}
\begin{minipage}{.49\linewidth}
\centering
  \includegraphics[height=1.5in]{fac-overview.png}
\end{minipage}
\caption{(L) An ill-typed \texttt{fac} function; (R) Dynamically witnessing the type error in \texttt{fac}.}
\label{fig:factorial}
\end{figure}

In this paper we propose a new approach that explains
static type errors by \emph{dynamically} witnessing
how an ill-typed program goes wrong.
%
We have developed \toolname, an interactive tool that uses
the source of the ill-typed function to automatically synthesize
the result on the right in Figure~\ref{fig:factorial}, which
shows how the recursive calls reduce to a configuration where
the program ``goes wrong'' --- \ie\ the @int@ value @1@ is to be
multiplied with the @bool@ value @true@.
We achieve this via three concrete contributions.

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:semantics}).
%
This problem is surprisingly tricky when we cannot rely on
static type information.
%
In particular, we must avoid the trap of \emph{spurious} inputs
that cause irrelevant problems that would be avoided by picking
values of a different, relevant type.
%
We solve this problem by developing a novel operational semantics
that corresponds to executing the program with \emph{holes} ---
values whose type is unknown.
%
Our semantics \emph{conservatively} instantiates holes with concrete
values; thereby dynamically inferring the type of the input
until the program goes wrong.
%
We prove that our procedure synthesizes \emph{general witnesses},
which means, intuitively, that if a witness is found for a given
ill-typed function, then, \emph{for all} input types, there exist
inhabitants that can make the function go wrong.

@@

>>>>

\begin{figure}[t]
\centering
\begin{minipage}{.49\linewidth}
\centering
\begin{code}
  let rec fac n =
    if n <= 0 then
      true
    else
      n * fac (n-1)
\end{code}
\vspace{3em}
% \end{minipage}
% % \hspace{0.3in}
% \begin{minipage}{.39\linewidth}
\centering
  \includegraphics[height=1.5in]{fac-overview.png}
\end{minipage}
\begin{minipage}{.49\linewidth}
\centering
\includegraphics[height=3in]{fac-long.png}
\end{minipage}
\caption{(TL) An ill-typed \texttt{fac} function; (BL) Dynamically witnessing the type error in \texttt{fac}, showing only function calls; (R) The same trace, fully expanded.}
\label{fig:factorial}
\end{figure}

In this paper we propose a new approach that explains
static type errors by \emph{dynamically} witnessing
how an ill-typed program goes wrong.
%
We have developed \toolname, an interactive tool that uses
the source of the ill-typed function to automatically synthesize
the result on the bottom-left in Figure~\ref{fig:factorial}, which
shows how the recursive calls reduce to a configuration where
the program ``goes wrong'' --- \ie\ the @int@ value @1@ is to be
multiplied with the @bool@ value @true@.
We achieve this via three concrete contributions.

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:semantics}).
%
This problem is surprisingly tricky when we cannot rely on
static type information.
%
In particular, we must avoid the trap of \emph{spurious} inputs
that cause irrelevant problems that would be avoided by picking
values of a different, relevant type.
%
We solve this problem by developing a novel operational semantics
that combines evaluation and type inference.
%
We execute the program with \emph{holes} --- values whose type is
unknown --- as the inputs.
% that corresponds to executing the program with \emph{holes} ---
% values whose type is unknown.
%
A hole remains abstract until the evaluation context tells us what type
it must have, for example the parameters to an addition operation must
both be integers.
%
Our semantics \emph{conservatively} instantiates holes with concrete
values; thereby dynamically inferring the type of the input
until the program goes wrong.
%
We prove that our procedure synthesizes \emph{general witnesses},
which means, intuitively, that if a witness is found for a given
ill-typed function, then, \emph{for all} input types, there exist
inhabitants that can make the function go wrong.

@@

We introduce a notion of \emph{jump-compressed} traces to abstract away
the irrelevant details of a computation.
%
A jump-compressed trace includes only big steps to the next function
call or return, for example the trace in the bottom-left of
Figure~\ref{fig:factorial} is a jump-compressed trace.
%
Jump-compressed traces are similar to stack traces in that both show a
sequence of function calls that lead to a crash, but the jump-compressed
trace also shows the return values of successful calls, which can be
useful in understanding why a particular path was taken.
